### 2020-06-26 docs/LaTex/Paper.tex
Completes docs

% \newpage
% \section{Background}
\newpage
\section{Technical description}
Like in the Introduction, the most proper way to present this technical description is to difference between software and hardware parts. However, is still interesting to present a simple and general diagram for whole designed solution.

\begin{figure}[htp]
  \centering
  \includegraphics[width=.73\textwidth]{../schemes/main_scheme_tbg.png}
  \caption{General view of this solution.}
  \label{fig:GenView}
\end{figure}

@@

\subsubsubsection{Signal conditioning}
This stage is focused on improve the signal acquisition, according to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option. \doubt{Figure \ref{fig:SignalCond} shows a possible schematics for this stage}

@@

By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as can be read also in the Pin Mapping section of the library \cmd{README.md}. As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.

@@

  // Read the analog value of the potmeter (0-1023)
  potVal = analogRead(potPin);

@@

So, as can be seen, in order to prepare the payload firstly the analog pin is stored in \cmd{potVal} variable, after this the function \cmd{lpp.reset()} is ued to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time.
%So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields. The working of this function is explained in \cite{MCCI_lmic_16bitencoder}.\footnote{A javascript decoder is also provided by the \href{https://github.com/mcci-catena/arduino-lmic\#uflt16-javascript-decoder}{\cmd{README.md}} but \doubt{it will be described in next sections.}}

These \cmd{lpp.*()} functions are part of the CayenneLPP library, which is included at line 
%This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{payloadPot} 16-bit integer is split into two bytes called \cmd{potLow} and \cmd{potHigh} ---using native \cmd{lowByte} and \cmd{highByte} functions. Finally these bytes are stored in the first two elements of initial \cmd{payload[3]} string declared at line 65, as the reader can see in example \ref{ej:potmeter_vars}, and the payload is prepared to send at the next possible time using the function \cmd{LMIC\_setTxData2()}.

It is important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup() \{\}} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}

\begin{lstlisting}[%
  % float=ht,
  language=cpp,
  firstnumber=250,
  caption={Channels setup.},
  label=ej:chan_setup
  ]
  LMIC_setupChannel(0, 868100000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(1, 868300000, DR_RANGE_MAP(DR_SF12, DR_SF7B), BAND_CENTI);  // g-band
  LMIC_setupChannel(2, 868500000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(3, 867100000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(4, 867300000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(5, 867500000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(6, 867700000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(7, 867900000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(8, 868800000, DR_RANGE_MAP(DR_FSK,  DR_FSK),  BAND_MILLI);  // g2-band
\end{lstlisting}

@@

Then it would be necessary to install the GPIO access library from Raspberry repositories, so performing \cmd{sudo\, apt\, install\, wiringpi}. After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi}

@@

>>>>

% \newpage
% \section{Background}
\newpage
\section{Technical description}
Like in the Introduction, the most proper way to present this technical description is to difference between software and hardware parts. However, is still interesting to present a simple and general diagram for whole designed solution. So as the reader can see in the figure \ref{fig:GenView} is shown the general diagram.

Firstly the TTN stack send through MQTT protocol the payload to FIWARE IoTagent-LoRaWAN generic enabler. Secondly, FIWARE generic enablers process that payload and store the contained information (also called \textit{context data}) along with a timestamp in the persisting data base (CrateDB), after this, the context data can be visualized in a web view provided by Grafena.

At this moment this solution requires to send manually the HTTP requests in order to communicate with FIWARE generic enablers manually. The ideal situation would be to develop a desktop or web application capable of provide any kind of graphical user interface to easily setup the two most important requirements to get the solution working as intended ---Create a new device using the IoTagent-LoRaWAN, create a subscription in Orion Context Broker which sends notifications to QuantumLeap.

\begin{figure}[htp]
  \centering
  \includegraphics[width=\textwidth]{../schemes/main_scheme_tbg.png}
  \caption{General view of this solution.}
  \label{fig:GenView}
\end{figure}

@@

\subsubsubsection{Signal conditioning}
This stage is focused on improve the signal acquisition, according to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option. \doubt{Figure \ref{fig:SignalCond} shows a possible schematics for this stage.}

It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead a variable resistor, this is because taking advantage of the Arduino power source, particularly the 5 \si{\volt} and GND pins.

@@

By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}. As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.

@@

  // Read the analog value of the potmeter (0-1023)
  potVal = analogRead(potPin);
  potVal = 100.0 * potVal / 1023;

@@

So, as can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time.
%So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields. The working of this function is explained in \cite{MCCI_lmic_16bitencoder}.\footnote{A javascript decoder is also provided by the \href{https://github.com/mcci-catena/arduino-lmic\#uflt16-javascript-decoder}{\cmd{README.md}} but \doubt{it will be described in next sections.}}

These \cmd{lpp.*()} functions are part of the CayenneLPP library, in particular, the \cmd{lpp.addAnalogInput()} function encodes the raw value as a number encoded in 2 bytes (16 bits signed) with a unit of 0.01, in 16 bits signed is possible to represent numbers from -32768 until 32767, so with a unit of 0.01, that amounts to a range from -327.68 to +327.67; that is because the sketch includes the line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L185}{185}, where the raw value \cmd{potVal} is scaled to a [0, 100.0] range. This range will fit in the range that can be encoded in CayenneLPP. 
%This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{payloadPot} 16-bit integer is split into two bytes called \cmd{potLow} and \cmd{potHigh} ---using native \cmd{lowByte} and \cmd{highByte} functions. Finally these bytes are stored in the first two elements of initial \cmd{payload[3]} string declared at line 65, as the reader can see in example \ref{ej:potmeter_vars}, and the payload is prepared to send at the next possible time using the function \cmd{LMIC\_setTxData2()}.

CayenneLPP is so importante because FIWARE will work with this data model, so any other third party encoding will not be admitted/decoded by the FIWARE LoRaWAN agent.

It is also important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup() \{\}} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}

\begin{lstlisting}[%
  % float=ht,
  language=cpp,
  firstnumber=251,
  caption={Channels setup.},
  label=ej:chan_setup
  ]
  LMIC_setupChannel(0, 868100000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(1, 868300000, DR_RANGE_MAP(DR_SF12, DR_SF7B), BAND_CENTI);  // g-band
  LMIC_setupChannel(2, 868500000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(3, 867100000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(4, 867300000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(5, 867500000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(6, 867700000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(7, 867900000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
  LMIC_setupChannel(8, 868800000, DR_RANGE_MAP(DR_FSK,  DR_FSK),  BAND_MILLI);  // g2-band
\end{lstlisting}

@@

Then it would be necessary to install the GPIO access library from Raspberry repositories, so performing \cmd{sudo\, apt\, install\, wiringpi} the package manager should install \cmd{wiringpi} providing all necessary libraries.

After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi} and compile it performing the command shown in example \ref{ej:compiling_dual_channel_pkg} ---Like will be explained below, it is also important to note where \cmd{git} will clone the repository and where the reader will compile it, because the location of the \cmd{dual\_chan\_pkt\_fwd} binary is so critical for that systemd service (also included in the GitHub repository) \cmd{dual\_chan\_pkt\_fwd.service} works properly.

@@

The core of the software part is deployed using Docker containers \cite{Docker_container}. These docker containers are deploying essentially the following components

\begin{itemize}
  \item \textbf{IoTagent-LoRaWAN}, which is in charge of forward the traffic from TTN stack to Orion Context Broker.
  \item \textbf{Orion Context Broker}, this is the \textbf{core generic enabler}, in fact as indicated, this is the only one required component to the solution be considered as \enquote{powered by FIWARE}. Besides this generic enabler requires a \textbf{Mongo database}.
  \item \textbf{QuantumLeap}, in order to make the context data persisting, this generic enabler is needed. Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the information will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persisting.
\end{itemize}
